Computing solution space properties of combinatorial optimization problems by generic programming tensor networks

### 2022-04-13 paper/paper.tex
Updates from Overleaf

\title{Computing solution space properties of combinatorial optimization problems by generic programming tensor networks
\thanks{\funding{...}}
}

\author{XXX\thanks{XXX 
  (\email{email}, \url{website}).}
\and YYY\thanks{yyyyy 
  (\email{yyyy}, \email{email}).}
}

@@

The corresponding tensor network contraction cab be defined as
\begin{equation}\label{eq:idp}
    P(G, \{x_1^{w_1}, x_{2}^{w_2}, \ldots,x_{|V|}^{w_{|V|}}\}) = \sum\limits_{s_1, s_2, \ldots, s_{|V|} = 0}^{1} \prod\limits_{v\in V} W(x_v^{w_v})_{s_v} \prod\limits_{(u,v) \in E} B_{s_u s_v},
\end{equation}
where the summation runs over all $2^{|V|}$ vertex configurations $\{s_1, s_{2}, \ldots,s_{|V|}\}$ and accumulates the product of tensor elements to the output $P$. 
% It is easy to see that an edge tensor element $B_{s_us_v}$ encodes the independent set constraint. It zeros out the product whenever both $u$ and $v$ are in the independent set.
A vertex tensor element $W(x_v^{w_v})_{s_v}$ contributes a multiplicative factor $x_v^{w_v}$ whenever $v$ is in the set.

@@

\section{Maximum independent sets and its counting}\label{sec:counting}
\subsection{Tropical algebra for finding the MIS size and counting MISs}
%The method we use to count the MISs is based on the following observations:
Let $x=\infty$, the independence polynomial in the previous section becomes
\begin{equation}
I(G, \infty) = \lim_{x\rightarrow \infty}\sum_{k=0}^{\alpha(G)}a_k x^k = a_{\alpha(G)} \infty^{\alpha(G)},
\end{equation}
where all terms except the one with the largest order vanish. We can thus replace the polynomial type $a = (a_0, a_1, \ldots, a_k)$ with a new type that has two fields: the largest exponent $k$ and its coefficient $a_k$.
From this, we can define a new algebra as
\begin{equation}
    \eqname{P1}
\begin{split}
    a_x\infty^x \oplus a_y\infty^y &= \begin{cases}
        (a_x + a_y)\infty^{\max(x,y)}, & x = y\\
        a_y\infty^{\max(x,y)}, & x < y\\
        a_x\infty^{\max(x,y)}, & x > y
    \end{cases}, \\
    a_x\infty^x \odot a_y\infty^y &= a_x a_y\infty^{x+y}\\
    \mymathbb{0} &= 0\infty^{-\infty}\\
    \mymathbb{1} &= 1\infty^{0}.
\end{split}%& \text{\rlap{(P1)}}
\label{eq:countingtropical}
\end{equation}
Here, we have generalized the previous polynomial to the Laurent polynomial to define the zero-element properly.
To implement this algebra programmatically, we create a data type with two fields $(x, a_x)$ to store the MIS size and its counting,
and define the above operations and constants correspondingly.
If one is only interested in finding the MIS size, one can drop the counting field.
The algebra of the exponents becomes the max-plus tropical algebra~\cite{Maclagan2015, Moore2011}:
\begin{equation}\eqname{T}
    \begin{split}
        x \oplus y &= \max(x,y)\\
        x \odot y &= x + y\\
        \mymathbb{0} &= -\infty\\
        \mymathbb{1} &= 0.
    \end{split}\label{eq:tropical}
\end{equation}
Algebra \Eq{eq:tropical} and \Eq{eq:countingtropical} are the same as those used in Liu et al.~\cite{Liu2021} to compute the spin glass ground state energy and its degeneracy.
For independent set calculations here, the vertex tensor and edge tensor becomes:
\begin{equation}
    W^{\rm T} = \left(\begin{matrix}
        \mymathbb{1} \\
        \infty
    \end{matrix}\right),   
    \qquad \qquad
        B^{\rm T} = \left(\begin{matrix}
        \mymathbb{1}  & \mymathbb{1} \\
        \mymathbb{1} & \mymathbb{0}
    \end{matrix}\right).
\end{equation}

@@

This set algebra can serve as the coefficients in \Eq{eq:polynomial} to enumerate independent sets of all different sizes, \Eq{eq:countingtropical} to enumerate all MISs, or \Eq{eq:max2poly} to enumerate all independent sets of sizes $\alpha(G)$ and $\alpha(G)-1$.
As long as the coefficients in a truncated polynomial are members of a commutative semiring, the polynomial itself is a commutative semiring.
For example, to enumerate only the MISs, we can define a combined element type $s_{k}\infty^k$, where the coefficients follow the algebra in \Eq{eq:set} and the exponents follow the max-plus tropical algebra.
The combined operations become: 
\begin{equation}
\eqname{P1+SN}
\begin{split}
    s_x\infty^x \oplus s_y\infty^y &= \begin{cases}
        (s_x \cup s_y)\infty^{\max(x,y)}, & x = y\\
        s_y\infty^{\max(x,y)}, & x < y\\
        s_x\infty^{\max(x,y)}, & x > y
    \end{cases},\\
    s_x\infty^x \odot s_y\infty^y &= \{\sigma \lor^\circ \tau | \sigma \in s_x, \tau \in s_y\}\infty^{x+y},\\
    \mymathbb{0} &= \{\}\infty^{-\infty},\\
    \mymathbb{1} &= \{0^{\otimes |V|}\}\infty^{0}. \label{eq:countingtropicalset}
\end{split}
\end{equation}
Clearly, the vertex tensor and edge tensor become
\begin{equation}
    W^{\rm P1+SN}(\{\boldsymbol{e}_{i}\}\infty^1) = \left(\begin{matrix}
        \mymathbb{1} \\
        \{\boldsymbol{e}_{i}\}\infty^1
    \end{matrix}\right),   
    \qquad \qquad
        B^{\rm P1+SN} = \left(\begin{matrix}
        \mymathbb{1}  & \mymathbb{1} \\
        \mymathbb{1} & \mymathbb{0}
    \end{matrix}\right).
\end{equation}
The enumeration of all MIS configurations corresponds to the contraction of this tensor network.
However, direct contraction might have significant space overheads for keeping too many intermediate states irrelevant to the final maximum independent sets.
We introduce the bounding technique in \App{sec:bounding} to avoid this issue.
One may also be interested in the more studied maximal independent sets~\cite{Bron1973, Eppstein2010, Johnson1988} enumeration.
It will be discussed in \App{sec:maximal} since it requires using a different tensor network structure.

@@

\section{Weighted graphs}\label{sec:weighted}
All the solution space properties and the corresponding algebra on unweighted graphs still hold for integer-weighted graphs, while for general weighted graphs, the independence polynomial is not well defined anymore.
For general weighted graphs, it is more useful to know the $k$ maximum weighted sets and their sizes.
They can be computed by the extended tropical algebra, which is a natural generalization of the max-plus tropical algebra:
\begin{equation}
\eqname{T$k$}
\begin{split}
    s \oplus t &= \texttt{largest}(s \cup t, k)\\
    s \odot t &= \texttt{largest}(\{a+b | a \in s, b\in t\}, k)\\
    \mymathbb{0} &= -\infty^{\otimes k}\\
    \mymathbb{1} &= -\infty^{\otimes k-1} \otimes 0
\end{split}\label{eq:ext-tropical}
\end{equation}
where $\texttt{largest}(s, k)$ means truncating the set $s$ by only keeping its $k$ largest values.
The computation of $s \odot t$ is a maximum sum combination problem that can be done in time $O(k\log(k))$.
An efficient algorithm for computing the maximum sum combination problem can be found in \App{sec:maxsum}.
The vertex tensor and edge tensor become
\begin{equation}
    W^{{\rm T}k}(-\infty^{\otimes k-1} \otimes 1, w_v) = \left(\begin{matrix}
        \mymathbb{1} \\
        -\infty^{\otimes k-1} \otimes w_v
    \end{matrix}\right),   
    \qquad
        B^{{\rm T}k} = \left(\begin{matrix}
        \mymathbb{1}  & \mymathbb{1} \\
        \mymathbb{1} & \mymathbb{0}
    \end{matrix}\right),
\end{equation}
where we have used the equation $(-\infty^{\otimes k-1} \otimes 1)^{w_v} = -\infty^{\otimes k-1} \otimes w_v$.
To find solutions corresponding to the largest $k$ sizes, one can combine the extended tropical algebra with the bit string algebra (\Eq{eq:singleconfig}).
Since the $\oplus$ operation of the configuration sampler will not be used in the combined algebra, the resulting configurations are deterministic and complete.

@@

\section*{Acknowledgments}
We would like to thank Pan Zhang for sharing his python code for optimizing contraction orders of a tensor network.
We acknowledge Sepehr Ebadi and Leo Zhou for coming up with many interesting questions about independent sets and their questions strongly motivated the development of this project.
We thank Benjamin Schiffer for providing helpful feedback on the writing of this manuscript.
We thank Chris Elord for helping us write the fastest matrix multiplication library for tropical numbers, TropicalGEMM.jl. % he is a man of speed!
We would also like to thank a number of open-source software developers, including Roger Luo, Time Besard, Edward Scheinerman, and Katharine Hyatt
for actively maintaining their packages and resolving related issues voluntarily.
\blue{funding information, need to mention the AWS}

@@

Let us characterize the independent set problem on graph $G=(V, E)$ as an energy model with two parts
\begin{equation}\label{eq:eng}
    \mathcal{E}(G, s) = -\sum_{i\in V} w_i s_i + \infty \sum_{\langle i,j\rangle \in E}s_i s_j
\end{equation}
where $s_i$ is a spin on vertex $i \in V$ and $w_i$ is an onsite energy term associated with it.
The first part corresponds to the negative independent set size and the second part describes the independence constraint, which corresponds to the Rydberg blockade~\cite{Pichler2018, Ebadi2022} in cold atom arrays or the repulsive force in hardcore lattice models~\cite{Dyre2016, Fernandes2007}.
The partition function is defined as
\begin{equation}\label{eq:partition}
    \begin{split}
    Z(G, \beta) = \sum_{s}e^{-\beta \mathcal{E}(G, s)} = \sum_{s\in \mathcal{I}(G)} e^{\beta \sum w_i s_i}\\
         = \sum_{k=0}^{\alpha(G)}a(k) e^{\beta k}  \qquad \quad (k = \sum w_i s_i)
    \end{split}
\end{equation}
where $\mathcal{I}(G)$ is the set of independent sets of graph $G$, $\alpha(G)$ is the absolute value of the minimum energy (maximum independent set size), $a(k)$ is the number of spin configurations with energy $-k$ (independent sets of size $k$).
The partition function can be expressed as a tensor network by placing a vertex tensor on each spin $i$
\begin{equation}
    W(e^{\beta w_i}) = \left(\begin{matrix}
        1 \\
        e^{\beta w_i}
    \end{matrix}\right),
\end{equation}
and an edge tensor on each bond
\begin{equation}
       B = \left(\begin{matrix}
        1  & 1\\
        1 & 0
    \end{matrix}\right),
\end{equation}
where the $0$ in the edge tensor comes from $e^{-\beta\infty}$ in the second term of \Eq{eq:eng}, which is the independence constraint.
By letting $x = e^{\beta}$, we get the tensor network for computing the independence polynomial as described by \Eq{eq:vertextensor} and \Eq{eq:edgetensor}.
If we further let $w_i=1$, the second line of \Eq{eq:partition} is equivalent to the independence polynomial.

@@

\section{Hard problems and tensor networks}\label{sec:otherproblems}
\subsection{Maximal independent sets and maximal cliques}\label{sec:maximal}
In this section, we focus the discussion on the maximal independent sets problem since finding maximal cliques of a graph is equivalent to finding the maximal independent sets of the complement of the graph.
Let $G=(V,E)$ be a graph, we denote the neighborhood of a vertex $v\in V$ as $N(v)$.
A maximal independent set $I_m$ is an independent set such that no $v \in V$ satisfies $I_m \cap (\{v\} \cup N[v])  = \emptyset$, i.e.\ an independent set that cannot become a larger one by adding a new vertex.
To characterize the maximal independence restriction, we defined a tensor on each $\{v\}\cup N(v)$ as
\begin{equation}\label{eq:maximal}
    T(x_v, w_v)_{s_1,s_2,\ldots,s_{|N(v)|},s_v} = \begin{cases}
        s_vx_v^{w_v} & s_1=s_2=\ldots=s_{|N(v)|}=0,\\
        1-s_v& \text{otherwise}.\\
    \end{cases}
\end{equation}
It means if none of the neighbours of $v$ are in $I_{m}$ ($ s_1=s_2=\ldots=s_{|N(v)|}=0$), then $v$ must be in $I_{m}$ and contribute a factor $x_{v}$, otherwise, if any of the neighbourhood vertices is in $I_{m}$, then $v$ cannot be in $I_{m}$.
For a vertex $v$ that has a degree 2, this tensor has the following form
\begin{equation}
    T(x_v, w_v)=\left(\begin{matrix}
    \left(\begin{matrix}
        ~~0 &~1 \\
        ~~1 &~1
    \end{matrix}\right)\\
    \left(\begin{matrix}
        x_v^{w_v} &0 \\
        0 &0
    \end{matrix}\right)
    \end{matrix}\right).
\end{equation}

@@

\subsection{Boolean satisfiability Problem}
The Boolean satisfiability problem is the problem of determining if there exists an assignment that satisfies a given Boolean formula.
One can specify a satisfiable problem in the conjunctive normal form (CNF), i.e.\ a conjunction of clauses (or disjunctions of Boolean literals).
To reduce the problem of solving a CNF to a tensor network contraction, we first map a Boolean literal $a$ and its negation $\neg a$ to the same degree of freedom $s_a \in \{0, 1\}$.
$s_a = 0$ stands for variable $a$ having value \texttt{false} while $s_a=1$ stands for having value \texttt{true}.
Then we map a clause to a tensor. For example, a $k$-th clause $\neg a \vee b \vee \neg c$ can be mapped to a tensor labeled by $(s_a, s_b, s_c)$.
\begin{equation}
C_{k} = \left(\begin{matrix}
\left(\begin{matrix}
x^{w_k} & x_{b}^{w_k} \\
x_a^{w_k} & x_{ab}^{w_k}
\end{matrix}\right) \\
\left(\begin{matrix}
x_{c}^{w_k} & x_{bc}^{w_k} \\
1_{ac} & x_{abc}^{w_k}
\end{matrix}\right)
\end{matrix}\right),
\end{equation}
where $w_k$ is a weight associated with a clause. There is only one entry $(s_a, s_b, s_c) = (1, 0, 1)$ that makes this clause unsatisfied.
Let $x^{w_k}_{\ldots} = x$, one can get a polynomial, in which the $k$-th coefficient gives the number of assignments that $k$ clauses are satisfied.

\subsection{Set packing}
Suppose one has a finite set $S$ and a list of subsets of $S$. Then, the set packing problem asks if some $k$ subsets in the list are pairwise disjoint.
It is the hypergraph generalization of the maximum independent set problem, where a set corresponds to a vertex and an element corresponds to a hyperedge.
To solve the set packing problem, we generalize the edge tensor in \Eq{eq:edgetensor} as
\begin{equation}
    B_{s_u,s_v,\ldots, s_w} = \begin{cases}
        1, & s_u+s_v+\ldots+s_w\leq 1,\\
        0, & \text{otherwise}.
    \end{cases}
\end{equation}

>>>>

\title{Computing solution space properties of combinatorial optimization problems by generic programming tensor networks
%\thanks{\funding{...}}
}

\author{
Jin-Guo Liu\footnote{JIN-GUO LIU and XUN GAO contributed equally to this paper.}\thanks{Department of Physics, Harvard University, Cambridge, Massachusetts 02138, USA; QuEra Computing Inc., 1284 Soldiers Field Road, Boston, MA, 02135, USA %\thanks{QuEra Computing Inc., 1284 Soldiers Field Road, Boston, MA, 02135, USA}
    (\email{jinguoliu@g.harvard.edu}).}
\and {Xun Gao}\thanks{Department of Physics, Harvard University, Cambridge, Massachusetts 02138, USA (\email{xungao@g.harvard.edu}).}
\and Madelyn Cain\thanks{Department of Physics, Harvard University, Cambridge, Massachusetts 02138, USA}
\and Mikhail D. Lukin\thanks{Department of Physics, Harvard University, Cambridge, Massachusetts 02138, USA}
\and Sheng-Tao Wang\thanks{QuEra Computing Inc., 1284 Soldiers Field Road, Boston, MA, 02135, USA}
}

@@

The corresponding tensor network contraction can be defined as
\begin{equation}\label{eq:idp}
    P(G, \{x_1^{w_1}, x_{2}^{w_2}, \ldots,x_{|V|}^{w_{|V|}}\}) = \sum\limits_{s_1, s_2, \ldots, s_{|V|} = 0}^{1} \prod\limits_{v\in V} W(x_v^{w_v})_{s_v} \prod\limits_{(u,v) \in E} B_{s_u s_v},
\end{equation}
where the summation runs over all $2^{|V|}$ vertex configurations $\{s_1, s_{2}, \ldots,s_{|V|}\}$ and accumulates the product of tensor elements to the output $P$. 
% It is easy to see that an edge tensor element $B_{s_us_v}$ encodes the independent set constraint. It zeros out the product whenever both $u$ and $v$ are in the independent set.
A vertex tensor element $W(x_v^{w_v})_{s_v}$ contributes a multiplicative factor $x_v^{w_v}$ whenever $v$ is in the set.

@@

\section{Maximum independent sets and its counting}\label{sec:counting}
\subsection{Tropical algebra for finding the MIS size and counting MISs}
%The method we use to count the MISs is based on the following observations:
Let $x=\infty$, the independence polynomial in the previous section becomes
\begin{equation}
I(G, \infty) = \lim_{x\rightarrow \infty}\sum_{k=0}^{\alpha(G)}a_k x^k = a_{\alpha(G)} \infty^{\alpha(G)},
\end{equation}
where all terms except the one with the largest order vanish. We can thus replace the polynomial type $a = (a_0, a_1, \ldots, a_k)$ with a new type that has two fields: the largest exponent $k$ and its coefficient $a_k$.
From this, we can define a new algebra as
\begin{equation}
    \eqname{P1}
\begin{split}
    a_x\infty^x \oplus a_y\infty^y &= \begin{cases}
        (a_x + a_y)\infty^{\max(x,y)}, & x = y\\
        a_y\infty^{\max(x,y)}, & x < y\\
        a_x\infty^{\max(x,y)}, & x > y
    \end{cases}, \\
    a_x\infty^x \odot a_y\infty^y &= a_x a_y\infty^{x+y}\\
    \mymathbb{0} &= 0\infty^{-\infty}\\
    \mymathbb{1} &= 1\infty^{0}.
\end{split}%& \text{\rlap{(P1)}}
\label{eq:countingtropical}
\end{equation}
Here, we have generalized the previous polynomial to the Laurent polynomial to define the zero-element properly.
To implement this algebra programmatically, we create a data type with two fields $(x, a_x)$ to store the MIS size and its counting,
and define the above operations and constants correspondingly.
If one is only interested in finding the MIS size, one can drop the counting field.
The algebra of the exponents becomes the max-plus tropical algebra~\cite{Maclagan2015, Moore2011}:
\begin{equation}\eqname{T}
    \begin{split}
        x \oplus y &= \max(x,y)\\
        x \odot y &= x + y\\
        \mymathbb{0} &= -\infty\\
        \mymathbb{1} &= 0.
    \end{split}\label{eq:tropical}
\end{equation}
Algebra \Eq{eq:tropical} and \Eq{eq:countingtropical} are the same as those used in Liu et al.~\cite{Liu2021} to compute the spin glass ground state energy and its degeneracy.
For independent set calculations here, the vertex tensor and edge tensor becomes:
\begin{equation}
    W^{\rm T}(w_v) = \left(\begin{matrix}
        w_v \\
        \infty
    \end{matrix}\right),   
    \qquad \qquad
        B^{\rm T} = \left(\begin{matrix}
        \mymathbb{1}  & \mymathbb{1} \\
        \mymathbb{1} & \mymathbb{0}
    \end{matrix}\right).
\end{equation}

@@

This set algebra can serve as the coefficients in \Eq{eq:polynomial} to enumerate independent sets of all different sizes, \Eq{eq:countingtropical} to enumerate all MISs, or \Eq{eq:max2poly} to enumerate all independent sets of sizes $\alpha(G)$ and $\alpha(G)-1$.
As long as the coefficients in a truncated polynomial are members of a commutative semiring, the polynomial itself is a commutative semiring.
For example, to enumerate only the MISs, we can define a combined element type $s_{k}\infty^k$, where the coefficients follow the algebra in \Eq{eq:set} and the exponents follow the max-plus tropical algebra.
The combined operations become: 
\begin{equation}
\eqname{P1+SN}
\begin{split}
    s_x\infty^x \oplus s_y\infty^y &= \begin{cases}
        (s_x \cup s_y)\infty^{\max(x,y)}, & x = y\\
        s_y\infty^{\max(x,y)}, & x < y\\
        s_x\infty^{\max(x,y)}, & x > y
    \end{cases},\\
    s_x\infty^x \odot s_y\infty^y &= \{\sigma \lor^\circ \tau | \sigma \in s_x, \tau \in s_y\}\infty^{x+y},\\
    \mymathbb{0} &= \{\}\infty^{-\infty},\\
    \mymathbb{1} &= \{0^{\otimes |V|}\}\infty^{0}. \label{eq:countingtropicalset}
\end{split}
\end{equation}
Clearly, the vertex tensor and edge tensor become
\begin{equation}
    W^{\rm P1+SN}(\{\boldsymbol{e}_{i}\}\infty^{w_v}) = \left(\begin{matrix}
        \mymathbb{1} \\
        \{\boldsymbol{e}_{i}\}\infty^{w_v}
    \end{matrix}\right),   
    \qquad \qquad
        B^{\rm P1+SN} = \left(\begin{matrix}
        \mymathbb{1}  & \mymathbb{1} \\
        \mymathbb{1} & \mymathbb{0}
    \end{matrix}\right).
\end{equation}
The enumeration of all MIS configurations corresponds to the contraction of this tensor network.
However, direct contraction might have significant space overheads for keeping too many intermediate states irrelevant to the final maximum independent sets.
We introduce the bounding technique in \App{sec:bounding} to avoid this issue.
One may also be interested in the more studied maximal independent sets~\cite{Bron1973, Eppstein2010, Johnson1988} enumeration.
It will be discussed in \App{sec:maximal} since it requires using a different tensor network structure.

@@

\section{Weighted graphs}\label{sec:weighted}
All the solution space properties and the corresponding algebra on unweighted graphs still hold for integer-weighted graphs, while for general weighted graphs, the independence polynomial is not well defined anymore.
For general weighted graphs, it is more useful to know the $k$ maximum weighted sets and their sizes.
They can be computed by the extended tropical algebra, which is a natural generalization of the max-plus tropical algebra:
\begin{equation}
\eqname{T$k$}
\begin{split}
    s \oplus t &= \texttt{largest}(s \cup t, k)\\
    s \odot t &= \texttt{largest}(\{a+b | a \in s, b\in t\}, k)\\
    \mymathbb{0} &= -\infty^{\otimes k}\\
    \mymathbb{1} &= -\infty^{\otimes k-1} \otimes 0
\end{split}\label{eq:ext-tropical}
\end{equation}
where $\texttt{largest}(s, k)$ means truncating the set $s$ by only keeping its $k$ largest values.
The computation of $s \odot t$ is a maximum sum combination problem that can be done in time $O(k\log(k))$.
An efficient algorithm for computing the maximum sum combination problem can be found in \App{sec:maxsum}.
The vertex tensor and edge tensor become
\begin{equation}
    W^{{\rm T}k}(-\infty^{\otimes k-1} \otimes w_v) = \left(\begin{matrix}
        \mymathbb{1} \\
        -\infty^{\otimes k-1} \otimes w_v
    \end{matrix}\right),   
    \qquad
        B^{{\rm T}k} = \left(\begin{matrix}
        \mymathbb{1}  & \mymathbb{1} \\
        \mymathbb{1} & \mymathbb{0}
    \end{matrix}\right),
\end{equation}
where we have used the equation $(-\infty^{\otimes k-1} \otimes 1)^{w_v} = -\infty^{\otimes k-1} \otimes w_v$.
To find solutions corresponding to the largest $k$ sizes, one can combine the extended tropical algebra with the bit string algebra (\Eq{eq:singleconfig}).
Since the $\oplus$ operation of the configuration sampler will not be used in the combined algebra, the resulting configurations are deterministic and complete.

@@

\section*{Acknowledgments}
We would like to thank Pan Zhang for sharing his python code for optimizing contraction orders of a tensor network.
We acknowledge Sepehr Ebadi and Leo Zhou for coming up with many interesting questions about independent sets and their questions strongly motivated the development of this project.
We thank Benjamin Schiffer for providing helpful feedback on the writing of this manuscript.
We thank Chris Elord for helping us write the fastest matrix multiplication library for tropical numbers, TropicalGEMM.jl. 
We would also like to thank a number of open-source software developers, including Roger Luo, Time Besard, Edward Scheinerman, and Katharine Hyatt
for actively maintaining their packages and resolving related issues voluntarily.
We acknowledge financial support from the DARPA ONISQ program (grant no.\ W911NF2010021), the Center for Ultracold Atoms, the National Science Foundation, the Vannevar Bush Faculty Fellowship, the U.S. Department of Energy (DE-SC0021013 and DOE Quantum Systems Accelerator Center (contract no.\ 7568717), 
%the Office of Naval Research, 
the Army Research Office MURI. We acknowledge the computation credits provided by Amazon Web Services for running the benchmarks and case studies. Jinguo Liu acknowledges funding support provided by QuEra Computing Inc.\ through a sponsored research program.

@@

Let us characterize the independent set problem on graph $G=(V, E)$ as an energy model with two parts
\begin{equation}\label{eq:eng}
    \mathcal{E}(G, s) = -\sum_{i\in V} w_i s_i + \infty \sum_{(i,j) \in E}s_i s_j
\end{equation}
where $s_i$ is a spin on vertex $i \in V$ and $w_i$ is an onsite energy term associated with it.
The first part corresponds to the negative independent set size and the second part describes the independence constraint, which corresponds to the Rydberg blockade~\cite{Pichler2018, Ebadi2022} in cold atom arrays or the repulsive force in hardcore lattice models~\cite{Dyre2016, Fernandes2007}.
The partition function is defined as
\begin{equation}\label{eq:partition}
    \begin{split}
    Z(G, \beta) = \sum_{s}e^{-\beta \mathcal{E}(G, s)} = \sum_{s\in \mathcal{I}(G)} e^{\beta \sum w_i s_i}\\
         = \sum_{k=0}^{\alpha(G)}a(k) e^{\beta k}  \qquad \quad (k = \sum w_i s_i)
    \end{split}
\end{equation}
where $\mathcal{I}(G)$ is the set of independent sets of graph $G$, $\alpha(G)$ is the absolute value of the minimum energy (maximum independent set size), $a(k)$ is the number of spin configurations with energy $-k$ (independent sets of size $k$).
The partition function can be expressed as a tensor network by placing a vertex tensor on each spin $i$
\begin{equation}
    W(e^{\beta w_i}) = \left(\begin{matrix}
        1 \\
        e^{\beta w_i}
    \end{matrix}\right),
\end{equation}
and an edge tensor on each bond
\begin{equation}
       B = \left(\begin{matrix}
        1  & 1\\
        1 & 0
    \end{matrix}\right),
\end{equation}
where the $0$ in the edge tensor comes from $e^{-\beta\infty}$ in the second term of \Eq{eq:eng}, which is the independence constraint.
By letting $x = e^{\beta}$, we get the tensor network for computing the independence polynomial as described by \Eq{eq:vertextensor} and \Eq{eq:edgetensor}.
If we further let $w_i=1$, the second line of \Eq{eq:partition} is equivalent to the independence polynomial.

@@

\section{Hard problems and tensor networks}\label{sec:otherproblems}
\subsection{Maximal independent sets and maximal cliques}\label{sec:maximal}
In this section, we focus the discussion on the maximal independent sets problem since finding maximal cliques of a graph is equivalent to finding the maximal independent sets of the complement of the graph.
Let $G=(V,E)$ be a graph, we denote the neighborhood of a vertex $v\in V$ as $N(v)$.
A maximal independent set $I_m$ is an independent set such that no $v \in V$ satisfies $I_m \cap (\{v\} \cup N[v])  = \emptyset$, i.e.\ an independent set that cannot become a larger one by adding a new vertex.
To characterize the maximal independence restriction, we defined a tensor on each $\{v\}\cup N(v)$ as
\begin{equation}\label{eq:maximal}
    T(x_v^{w_v})_{s_1,s_2,\ldots,s_{|N(v)|},s_v} = \begin{cases}
        s_vx_v^{w_v} & s_1=s_2=\ldots=s_{|N(v)|}=0,\\
        1-s_v& \text{otherwise}.\\
    \end{cases}
\end{equation}
It means if none of the neighbours of $v$ are in $I_{m}$ ($ s_1=s_2=\ldots=s_{|N(v)|}=0$), then $v$ must be in $I_{m}$ and contribute a factor $x_{v}$, otherwise, if any of the neighbourhood vertices is in $I_{m}$, then $v$ cannot be in $I_{m}$.
For a vertex $v$ that has a degree 2, the tensor has the following form
\begin{equation}
    T(x_v^{w_v})=\left(\begin{matrix}
    \left(\begin{matrix}
        ~~0 &~1 \\
        ~~1 &~1
    \end{matrix}\right)\\
    \left(\begin{matrix}
        x_v^{w_v} &0 \\
        0 &0
    \end{matrix}\right)
    \end{matrix}\right).
\end{equation}

@@

\subsection{Boolean satisfiability Problem}
The Boolean satisfiability problem is the problem of determining if there exists an assignment that satisfies a given Boolean formula.
One can specify a satisfiable problem in the conjunctive normal form (CNF), i.e.\ a conjunction of clauses (or disjunctions of Boolean literals).
To reduce the problem of solving a CNF to a tensor network contraction, we first map a Boolean literal $a$ and its negation $\neg a$ to the same degree of freedom $s_a \in \{0, 1\}$.
$s_a = 0$ stands for variable $a$ having value \texttt{false} while $s_a=1$ stands for having value \texttt{true}.
Then we map a clause to a tensor. For example, a $k$-th clause $\neg a \vee b \vee \neg c$ can be mapped to a tensor labeled by $(s_a, s_b, s_c)$.
\begin{equation}
C_{k} = \left(\begin{matrix}
\left(\begin{matrix}
x^{w_k} & x_{b}^{w_k} \\
x_a^{w_k} & x_{ab}^{w_k}
\end{matrix}\right) \\
\left(\begin{matrix}
x_{c}^{w_k} & x_{bc}^{w_k} \\
1 & x_{abc}^{w_k}
\end{matrix}\right)
\end{matrix}\right),
\end{equation}
where $w_k$ is a weight associated with a clause. There is only one entry $(s_a, s_b, s_c) = (1, 0, 1)$ that makes this clause unsatisfied.
Let $x^{w_k}_{\ldots} = x$, one can get a polynomial, in which the $k$-th coefficient gives the number of assignments that $k$ clauses are satisfied.

\subsection{Set packing}
Suppose one has a finite set $S$ and a list of subsets of $S$. Then, the set packing problem asks if some $k$ subsets in the list are pairwise disjoint.
It is the hypergraph generalization of the independent set problem, where a set corresponds to a vertex and an element corresponds to a hyperedge.
The generic tensor network for the set packing problem has a similar form as that for the independent set problem, where the vertex tensor is same as \Eq{eq:vertextensor}, while the edge tensor generalizes \Eq{eq:edgetensor} to higher dimensions
\begin{equation}
    B_{s_u,s_v,\ldots, s_w} = \begin{cases}
        1, & s_u+s_v+\ldots+s_w\leq 1,\\
        0, & \text{otherwise}.
    \end{cases}
\end{equation}
